---
title: "sf test: geometry coercion"
author: Tiernan Martin
output: html_notebook
date: November 7, 2016
---

This is a set of tests related to a [Github issue](https://github.com/edzer/sfr/issues/49) for the `sf` package. 


```{r setup}
library(tidyverse)
library(sf)
```

### Make test objects
Make a polygon, a multipolygon, and two dataframes that can be combined with the geometries to create simple feature objects.
```{r make-test-objs, echo=TRUE}

# Polygon geometry: pol

p1 <- rbind(c(0,0), c(1,0), c(3,2), c(2,4), c(1,4), c(0,0))
p2 <- rbind(c(1,1), c(1,2), c(2,2), c(1,1))
pol <-st_polygon(list(p1,p2))
# 
# Multi-polygon geometry: mpol

p3 <- rbind(c(3,0), c(4,0), c(4,1), c(3,1), c(3,0))
p4 <- rbind(c(3.3,0.3), c(3.8,0.3), c(3.8,0.8), c(3.3,0.8), c(3.3,0.3))[5:1,]
p5 <- rbind(c(3,3), c(4,2), c(4,3), c(3,3))
mpol <- st_multipolygon(list(list(p1,p2), list(p3,p4), list(p5)))
# 
# sf object w/ polygon: sf_pol

df1 <- 
        data.frame(VAR1 = 'foo',VAR2 = 'bar') %>% 
        mutate(geom = st_sfc(pol))
sf_pol <- st_as_sf(df1) 
# 
# sf object w/ multi-polygon: sf_mpol

df2 <- 
        data.frame(VAR1 = 'foo',VAR2 = 'bar') %>% 
        mutate(geom = st_sfc(mpol))
sf_mpol <- st_as_sf(df2)


# combined sf's, geometry type: GEOMETRY

sf_geometry <- rbind(sf_pol,sf_mpol)

sf_geometry


```



### Coercion 1: all geoms to `GEOMETRYCOLLECTION`
As mentioned in link, most ogr drivers cannot accept mixed-type geometries. Converting all geometries to `GEOMETRYCOLLECTION`s makes it possible to save these `sf` objects with `st_write()`.
```{r coercion-test-1, echo=TRUE}

coerce_to_geometrycollection <- function(x, crs = NULL){
        
        crs <- {
                if(is.null(crs)){
                        st_crs(x)
                }else{
                        crs
                } 
        }
        
        geom <- st_geometrycollection() %>% st_sfc()
        
        for(i in 1:nrow(x)){
                geom[[i]] <- st_geometry(x)[[i]] %>% list %>% st_geometrycollection()
        }
        
        x$geom <- geom
        st_crs(x) <- crs
        return(x)
        
}

sf_geometry %>% coerce_to_geometrycollection()

coerce_to_geom <- function(x, crs = NULL){
        
        crs <- {
                if(is.null(crs)){
                        st_crs(x)
                }else{
                        crs
                } 
        }
        
        geom <-  st_sfc()
        
        for(i in 1:nrow(x)){
                
                geom[[i]] <- {
                        if(
                                st_geometry(x)[[i]] %>% summary() %>% .[1,3] == "list"
                        ){
                                st_geometry(x)[[i]] %>% unlist(recursive = FALSE) %>% st_multipolygon()
                        }else{
                                st_geometry(x)[[i]] %>% unlist(recursive = FALSE) %>% st_polygon()
                        }
                }
        }
        
        x$geom <- geom
        st_crs(x) <- crs
        return(x)
        
}

sf_geometry %>% coerce_to_geom()

st_geometry(sf_geometry)[[1]] %>% unlist(recursive = FALSE) %>% summary()




sf_geometry %>% 
        coerce_to_geometrycollection() %>% 
        st_write(dsn = 'test.gpkg',driver = 'GPKG')

```


### Coercion 2: `GEOMETRYCOLLECTION` back to original geometries

```{r coercion-test-2, echo=TRUE}

test <- suppressWarnings(st_read('test.gpkg',quiet = TRUE)) 
test

coerce_from_geometrycollection <- function(x, crs = NULL){
        
        crs <- {
                if(is.null(crs)){
                        st_crs(x)
                }else{
                        crs
                } 
        }
        
        geom <- st_geometrycollection() %>% st_sfc()
        
        for(i in 1:nrow(x)){
                
                geom[[i]] <- {
                        if(
                                st_geometry(x)[[i]] %>% unlist(recursive = FALSE) %>% summary() %>% .[1,3] == "list"
                        ){
                                st_geometry(x)[[i]] %>% unlist(recursive = FALSE) %>% st_multipolygon()
                        }else{
                                st_geometry(x)[[i]] %>% unlist(recursive = FALSE) %>% st_polygon()
                        }
                }
        }
        
        x$geom <- geom
        st_crs(x) <- crs
        return(x)
        
}

test %>% coerce_from_geometrycollection()

```

