---
title: "sp × tidyverse test: rbind"
author: Tiernan Martin
output: html_notebook
date: November 5, 2016
---

This is a set of tests related to a [Github issue](https://github.com/edzer/sfr/issues/49) for the `sf` package. 


```{r setup}
library(tidyverse)
library(sf)
```

### Make test objects
Make a polygon, a multipolygon, and two dataframes that can be combined with the geometries to create simple feature objects.
```{r make-test-objs, echo=TRUE}

# Polygon geometry: pol

p1 <- rbind(c(0,0), c(1,0), c(3,2), c(2,4), c(1,4), c(0,0))
p2 <- rbind(c(1,1), c(1,2), c(2,2), c(1,1))
pol <-st_polygon(list(p1,p2))
# 
# Multi-polygon geometry: mpol

p3 <- rbind(c(3,0), c(4,0), c(4,1), c(3,1), c(3,0))
p4 <- rbind(c(3.3,0.3), c(3.8,0.3), c(3.8,0.8), c(3.3,0.8), c(3.3,0.3))[5:1,]
p5 <- rbind(c(3,3), c(4,2), c(4,3), c(3,3))
mpol <- st_multipolygon(list(list(p1,p2), list(p3,p4), list(p5)))
# 
# sf object w/ polygon: sf_pol

df1 <- 
        data.frame(VAR1 = 'foo',VAR2 = 'bar') %>% 
        mutate(geom = st_sfc(pol))
sf_pol <- st_as_sf(df1) 
# 
# sf object w/ multi-polygon: sf_mpol

df2 <- 
        data.frame(VAR1 = 'foo',VAR2 = 'bar') %>% 
        mutate(geom = st_sfc(mpol))
sf_mpol <- st_as_sf(df2)
```



### Test 1: rbind two sf objects with different geometry classes
```{r rbind-test, echo=TRUE}

# Check the sf object metadata: sf_pol
sf_pol

# Check the sf object metadata: sf_mpol
sf_mpol

# Attempt to combine the object with `rbind()`

sf_both <- rbind(sf_pol,sf_mpol)
sf_both

```

Looking at the metadata above, I'm surprised to see `geometry type: POLYGON`. Based on the `sf` [package vignette](https://edzer.github.io/sfr/articles/sfr.html#mixed-geometry-types), I would have guessed that the geometry of this new simple feature would be of the `GEOMETRY` class (but perhaps that was wishful thinking).

Let's see what happens when we plot this object:

```{r rbind-plot-test, echo=TRUE}
try(plot(sf_both))

```

It seems that `plot()` cannot handle this mixed-type `sfc`. It would be ideal if `rbind()` could detect the presence of different types of `sfg`'s and coerce them into a `GEOMETRY` `sfc`.

Otherwise, users need to find work-arounds. The following is one way to manually build the `GEOMETRY` `sfc` by looping over the two `sfg`'s:

```{r rbind-make-sfc-test,echo=TRUE}
sfc_new <- st_sfc()
        for(i in seq_along(st_geometry(sf_pol))){
                 
                 geom <- st_geometry(sf_pol)[[i]]
                 sfc_new[[i]] <- geom
        }
        for(i in seq_along(st_geometry(sf_mpol))){
                 
                 geom <- st_geometry(sf_mpol)[[i]]
                 sfc_new[[i + nrow(sf_pol)]] <- geom
                 
        }

sf_both_new <- 
        rbind(sf_pol,
              sf_mpol) %>% 
        mutate(geom = sfc_new)

sf_both_new
try(plot(sf_both_new))

```
So that work-around returns the desired results, but it wouldn't scale very well to accomodate the merger of many `sf` objects and it certainly breaks the continuity and readability of a `%>%` workflow (see princples 2 and 4 of the [Tidy Tools Manifesto](https://mran.microsoft.com/web/packages/tidyverse/vignettes/manifesto.html)).

### Test 2: bind sf objects with non-matching columns _and_ different geometry types

```{r bind_rows-test, echo = TRUE}

# sf object w/ polygon: sf_pol2
df3 <- data.frame(VAR1 = 'foo',VAR2 = 'bar') 
df3$geom <- st_sfc(pol)
sf_pol2 <- st_as_sf(df1) 
# 
# sf object w/ multi-polygon: sf_mpol2 (new column: VAR3)
df4 <- data.frame(VAR1 = 'foo',VAR3 = 'bar') 
df4$geom <- st_sfc(mpol)
sf_mpol2 <- st_as_sf(df4)
# 
# Try rbind()
sf_pol2;sf_mpol2

try(rbind(sf_pol2,sf_mpol2))
# 
# Try dplyr::bind_rows()
# NOTE: this throws an error, even with the help of try() 
# 
# try(dplyr::bind_rows(sf_pol2,sf_mpol2))
# 
# Work-around

df_both_new2 <- dplyr::bind_rows(sf_pol2[,1:2],
                                 sf_mpol2[,1:2])

df_both_new2$geom <- sfc_new

sf_both_new2 <- st_sf(df_both_new2)

sf_both_new2; try(plot(sf_both_new2))


```

`dplyr::bind_rows()` fails because it doesn't recognize the `sf` class. However, stripping the `sf` objects of their `sfc`'s, passing them to `bind_rows()`, and then adding the manually-built `GEOMETRY` `sfc` returns the desired result.


### Thoughts
Perhaps these tests make the case for the following enhancements to `sf`:

  1. an `sf`-compatible version of `dplyr::bind_rows()` (provided it's even possible - see Hadley's [comment](https://github.com/edzer/sfr/issues/49#issuecomment-258555275))
  2. an `sf` function for combining `sfc`'s with different geometry types into a `sfc` with `GEOMETRY` type `sfg`'s