---
df_print: tibble
output:
  html_notebook:
    code_folding: hide
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

```{r misc-setup, echo = FALSE, warning=FALSE,message=FALSE,comment=FALSE}
library(plyr)
library(knitr)
library(rprojroot)
library(rgdal)
library(sp)
library(rgeos)
library(tigris)
library(leaflet)
library(ggthemes)
library(magrittr)
library(stringr)
library(downloader)
library(webshot)
library(htmltools)
library(gplots)
library(ggmap)
library(shiny)
library(htmlwidgets)
library(readxl)
library(acs)
library(RColorBrewer)
library(tidyverse)
library(miscgis)
library(operator.tools)
library(sf)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE)
```

### Seattle Boundary

```{r misc-sea-bound, fig.cap="Seattle\'s geographic boundary"}

if(!file.exists(root_file('1-data/4-interim/sea-sp.gpkg'))){
        tigris::places(state = "WA") %>%
        tigris::filter_place(place = "Seattle") %>%
        spTransform(CRSobj = crs_proj) %>% 
        writeOGR(dsn = root_file('1-data/4-interim/sea-sp.gpkg'),
                 layer = "sea_sp", driver = "GPKG", verbose = FALSE, 
                 overwrite_layer = TRUE)
}

sea_sp <- readOGR(dsn = root_file('1-data/4-interim/sea-sp.gpkg'),layer = "sea_sp",verbose = FALSE) %>% spTransform(crs_proj)

green <- miscgis::miscgis_pals$tableau_cat[["green"]]

myLfltGrey() %>% addPolygons(data = sea_sp,color = green,opacity = 1,fillColor = green,fillOpacity = .5)

```


### Puget Sound Waterbodies (near Seattle)
These are useful for "clipping" census geographies whose boundaries extend into waterbodies.

```{r misc-sea-water, fig.cap="Waterbodies near Seattle"}

if(!file.exists(root_file('1-data/4-interim/wtr-sp.gpkg'))){
        fp_wtr <- root_file('1-data/3-external/NHDMajor.gdb')

# check if the file already exists, if not then download it
if(!file.exists(fp_wtr)){
        
        url <- "ftp://www.ecy.wa.gov/gis_a/inlandWaters/NHD/NHDmajor.gdb.zip" # save the URL for the waterbodies data
        
        temp <- tempfile() # create a temporary file to hold the compressed download
        
        download(url, dest = temp, mode="wb") # download the file
        
        unzip (temp, exdir = root_file('1-data/3-external/')) # extract the ESRI geodatabase file to a project folder
}

wtr_sp <-
        suppressWarnings(readOGR(dsn = fp_wtr,      # create a waterbodies shape
                layer = "NHD_MajorWaterbodies",verbose = FALSE,pointDropZ = TRUE)) %>%
        gBuffer(byid=TRUE, width=0) %>% # clean up self-intersecting polygons
        spTransform(CRSobj = crs_proj)  # transform the projection to match the project projection
        

intersect <- gIntersects(sea_sp,wtr_sp,byid = TRUE) %>% as.vector()

wtr_sp <- wtr_sp[intersect,] %>%
        spTransform(CRSobj = crs_proj)



writeOGR(obj = wtr_sp, dsn = root_file('1-data/4-interim/wtr-sp.gpkg'), 
         layer = "wtr_sp", driver = "GPKG", verbose = FALSE, 
         overwrite_layer = TRUE)
}

blue <- miscgis::miscgis_pals$tableau_cat[["blue"]]

wtr_sp <- readOGR(dsn = root_file('1-data/4-interim/wtr-sp.gpkg'),
                  layer = 'wtr_sp',
                  verbose = FALSE,
                  stringsAsFactors = FALSE) %>% spTransform(crs_proj)

myLfltGrey() %>% 
        addPolygons(data = wtr_sp,color = blue, opacity = 1, 
                    weight = .5, fillColor = blue,fillOpacity = .5)        

```

### Seattle Land
Seattle's boundary with waterbodies removed.

```{r misc-sea-land, fig.cap="Seattle without waterbodies"}

wtr_union_sp <- wtr_sp %>% gUnaryUnion()

if(!file.exists(root_file('1-data/4-interim/sea-lnd-sp.gpkg'))){


sea_lnd_sp <- miscgis::spClip(sea_sp,wtr_union_sp)

# Save the object
sea_lnd_sp %>% writeOGR(dsn = root_file('1-data/4-interim/sea-lnd-sp.gpkg'),
                        layer = 'sea_lnd_sp',
                        driver = 'GPKG',
                        verbose = FALSE)
        
}else{
        sea_lnd_sp <- readOGR(dsn = root_file('1-data/4-interim/sea-lnd-sp.gpkg'),
                        layer = 'sea_lnd_sp',
                        verbose = FALSE) %>% spTransform(crs_proj)
}


myLfltGrey() %>% addPolygons(data = sea_lnd_sp,color = green,opacity = 1,fillColor = green,fillOpacity = .5)

```


### Tracts in Seattle CCD

#### Intro
The Census tracts do not perfectly align with City's official boundary. As noted in the in a map produced by Seattle DPD (link below), Tracts 260.01, 264 and 265 extend beyond the boundary. The process below uses the Census's list of tracts 'within or partially-within' the city, which can be access using the [American Factfinder Tool](https://factfinder.census.gov/faces/nav/jsf/pages/index.xhtml). The process for accessing this data is explained in [here](http://www.seattle.gov/dpd/cs/groups/pan/@pan/documents/web_informational/dpdd017051.pdf).

```{r misc-tracts-sea}

if(!exists('tr_sea_sp')){
        
        if(!file.exists(root_file('1-data/4-interim/tr-sea-sp.gpkg'))){
        
        # Make a vector of the Seattle tract GEOIDs
        tr_sea_geoid6 <- read_csv(root_file("/1-data/3-external/manual/DEC_10_SF1_H1_with_ann.csv"),
                                 col_types = cols(Geography = col_skip(), 
                                                  Id = col_skip(), 
                                                  Id2 = col_character(),
                                                  Total = col_skip()), 
                                 skip = 1) %>%
                        mutate(GEOID6 = str_sub(Id2, start= -6)) %>%
                        select(GEOID6) %>%
                        filter(GEOID6 %!in% "990100") %>% # this tract is the water around Seattle (population = 0)
                        unlist(use.names = FALSE)
        
        # Download all census tracts in King County
        
        tr <- tigris::tracts(state = 53, county = 033) %>%
                                        spTransform(CRSobj = crs_proj)
        
        # Subset polygons (only Seattle) 
        
        tr@data %<>%  mutate(GEOID6 = str_sub(GEOID, start= -6))
        
        tr %<>% 
                subset(GEOID6 %in% tr_sea_geoid6) 
         
        tr %>% 
                miscgis::spClip(wtr_union_sp) %>% 
                writeOGR(dsn = root_file('1-data/4-interim/tr-sea-sp.gpkg'),
                         layer = 'tr_sea_sp',
                         driver = 'GPKG',
                         verbose = FALSE)
}
        
        
        tr_sea_sp <- readOGR(dsn = root_file('1-data/4-interim/tr-sea-sp.gpkg'),
                         layer = 'tr_sea_sp',
                         verbose = FALSE,
                         stringsAsFactors = FALSE) %>% spTransform(crs_proj)
}

myLfltGrey() %>% addPolygons(data = tr_sea_sp,
                             smoothFactor = 0,
                             fillOpacity = .25,
                             fillColor = green,
                             opacity = 1, 
                             color = green, 
                             weight = 1)


```
