---
df_print: tibble
output:
  html_notebook:
    code_folding: hide
  pdf_document:
    keep_tex: yes
always_allow_html: yes
---

```{r lmtd-eng-setup, echo = FALSE, warning=FALSE,message=FALSE,comment=FALSE}
library(plyr)
library(knitr)
library(rprojroot)
library(rgdal)
library(sp)
library(rgeos)
library(tigris)
library(leaflet)
library(ggthemes)
library(magrittr)
library(stringr)
library(downloader)
library(webshot)
library(htmltools)
library(gplots)
library(ggmap)
library(shiny)
library(htmlwidgets)
library(readxl)
library(acs)
library(RColorBrewer)
library(tidyverse)
library(miscgis)
library(operator.tools)
library(ggiraph)
library(leaflet.extras)
library(viridisLite)
library(sf)
root <- rprojroot::is_rstudio_project
root_file <- root$make_fix_file()
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, comment=FALSE)
green <- miscgis::miscgis_pals$tableau_cat[["green"]]
blue <- miscgis::miscgis_pals$tableau_cat[["blue"]]
orange <- miscgis::miscgis_pals$tableau_cat[["orange"]]
red <- miscgis::miscgis_pals$tableau_cat[["red"]]
teal <- miscgis::miscgis_pals$tableau_cat[["teal"]]
pal_rgb_4 <- miscgis_pals$tableau_cat[c("red","gold","green","blue")] %>% unlist %>% palette()
kc_tr_acs <- acs::geo.make(state = "WA",county = "King", tract = "*")

```

### ACS Data
Linguistic isolation processing steps:
  1. 
```{r lmtd-eng-acs}

# Setup the ACS geographies: Seattle (tracts), YCC (tracts), Seattle (place)
# -----------------------------------------------------------------------------

ycc_geo_acs <- read_rds(path = root_file('1-data/4-interim/ycc-geo-acs.rds'))

if(!file.exists(root_file('1-data/4-interim/lmtd-eng-orig-acs.rds'))){
        
        tbl <- "B16002"  # census table code
        
        acs.fetch(endyear = 2014,
                  geography = ycc_geo_acs,
                  table.number = tbl) %>%
                write_rds(root_file('1-data/4-interim/lmtd-eng-orig-acs.rds'))
        
}

lmtd_eng_orig_acs <- read_rds(root_file('1-data/4-interim/lmtd-eng-orig-acs.rds'))


# Process and save the data
# -----------------------------------------------------------------------------

if(!file.exists(root_file('1-data/4-interim/lmtd-eng-acs.rds'))){
        lmtd_eng_acs1 <- lmtd_eng_orig_acs
        
        guide <- 
                data.frame(terse = acs.colnames(lmtd_eng_acs1),
                       pretty = acs.colnames(acs.fetch(endyear = 2014, 
                          geography = geo.make(us = TRUE), 
                          table.number = tbl,col.names = 'pretty')))
        # Pull out the specific columns that are important
        
        get_col <- function(pattern){
                guide %>% filter(str_detect(pretty,pattern)) %>% select(terse) %>% unlist(use.names = FALSE) %>% as.character()
        }
        
        total <- get_col('Total')
        
        lmtd_eng <- get_col('Limited\\sEnglish\\sspeaking\\shousehold')
        
        
        # 1. Isolate the total column: 'TOTAL`

        acs_col_total <- lmtd_eng_acs1[,total]
        acs.colnames(acs_col_total) <- 'TOTAL'
        
        
        # 2. Combine all limited english speaking households: `LMTD_ENG`
        
        acs_col_lmtd <- apply(X = lmtd_eng_acs1[,lmtd_eng],
                                      MARGIN = 2,
                                      FUN = sum)
        acs.colnames(acs_col_lmtd) <- 'LMTD_ENG'
        
        # 3. Calculate the percentage of households paying 30% or more of their income on housing: `LMTD_ENG / TOTAL`
        
        acs_col_pct_lmtd_eng <- cbind(acs_col_total,acs_col_lmtd)
        acs_col_pct_lmtd_eng <-  
                acs::apply(X = acs_col_pct_lmtd_eng[,'LMTD_ENG'],
                           MARGIN = 1,
                           FUN = divide.acs,
                           denominator = acs_col_pct_lmtd_eng[,"TOTAL"],
                           method = "proportion",
                           verbose = FALSE)
        acs.colnames(acs_col_pct_lmtd_eng) <- 'PCT_LMTD_ENG'
        
# Save the object
# -----------------------------------------------------------------------------
        
        acs_col_pct_lmtd_eng %>% write_rds(root_file('1-data/4-interim/lmtd-eng-acs.rds'))
        
}

lmtd_eng_acs <- read_rds(root_file('1-data/4-interim/lmtd-eng-acs.rds'))


# Plot the data
# -----------------------------------------------------------------------------

plot(lmtd_eng_acs[order(estimate(lmtd_eng_acs)),1])


```

### Simple Feature object

```{r lmtd-eng-sf}

# Read in the `sf` object and re-order the records to by `JOIN_ID` 
# -----------------------------------------------------------------------------

ycc_all_sf <- 
        read_rds(root_file('1-data/4-interim/ycc-all-sf.rds'))

ycc_all_sf_ordered <- ycc_all_sf %>% arrange(JOIN_ID) %>% select(-JOIN_ID)

#
# Convert the `acs` object into a dataframe, join to sf object
# -----------------------------------------------------------------------------

ycc_lmtd_eng_sf <- 
        data.frame(
                geography(lmtd_eng_acs)["tract"],
                estimate(lmtd_eng_acs), 
                1.645 * standard.error(lmtd_eng_acs)) %>% 
        `colnames<-`(., c("JOIN_ID", "LMTD_ENG_PCT_EST","LMTD_ENG_PCT_MOE")) %>% 
        mutate(JOIN_ID = if_else(is.na(JOIN_ID),'SEA',JOIN_ID),
               UPPER = LMTD_ENG_PCT_EST + LMTD_ENG_PCT_MOE, 
               LOWER = LMTD_ENG_PCT_EST - LMTD_ENG_PCT_MOE, 
               LMTD_ENG_PCT_UPPER = if_else(UPPER > 1, 1, UPPER), 
               LMTD_ENG_PCT_LOWER = if_else(LOWER < 0, 0, LOWER)) %>% 
        select(JOIN_ID,everything(),-UPPER,-LOWER) %>% 
        arrange(JOIN_ID) %>% 
        bind_cols(ycc_all_sf_ordered) %>% 
        st_sf()

#
# Save the object
# -----------------------------------------------------------------------------
ycc_lmtd_eng_sf %>% write_rds(root_file('1-data/4-interim/lmtd-eng-sf.rds'))

```


### Dotplot (with variance)
Usage note: hovering a mouse cursor over a point reveal the name of the geography (e.g., a census tract ID or a neighborhood's name)

Notice the large error for Tract 53.02 (University District) - this reflects the large student population which drives down the census sample size.
```{r dotplot-lmtd-eng-sf}

# Show dotplot with variance
# -----------------------------------------------------------------------------

view_lmtd_eng_dotplot <- function() {
        
        df <- ycc_lmtd_eng_sf %>% select(-geom) %>% as.data.frame()
        
        gg <- ggplot(df, aes(x = reorder(JOIN_ID, 
                                             LMTD_ENG_PCT_EST), y = LMTD_ENG_PCT_EST))
        gg <- gg + geom_linerange(aes(ymin = LMTD_ENG_PCT_LOWER, 
                                      ymax = LMTD_ENG_PCT_UPPER), size = 1.5, alpha = 0.25)
        gg <- gg + geom_point_interactive(aes(tooltip = JOIN_ID), 
                                          size = 3, shape = 21, color = "white", fill = "grey30")
        gg <- gg + scale_x_discrete(expand = c(0.01, 
                                               0))
        gg <- gg + scale_y_continuous(labels = scales::percent)
        gg <- gg + coord_flip()
        gg <- gg + theme_minimal()
        gg <- gg + theme(panel.grid.major.y = element_blank(), 
                         panel.grid.minor.y = element_blank(), 
                         panel.grid.major.x = element_line(linetype = 3, 
                                                           color = "grey30"), 
                         panel.grid.minor.x = element_blank(), 
                         axis.text.y = element_blank())
        gg <- gg + labs(x = NULL, y = NULL, title = "Limited-English Households (%)", 
                        subtitle = "Census Tracts", 
                        caption = "Source: ACS Table B25106, Five-Year Estimates (2010-2014)")
        gg
}

view_lmtd_eng_dotplot()

# view_lmtd_eng_dotplot_int <- function() {
#         ggiraph(code = {
#                 print(view_lmtd_eng_dotplot())
#         }, width = 0.66, tooltip_extra_css = "padding:2px;background:rgba(70,70,70,0.1);color:black;border-radius:2px 2px 2px 2px;", 
#         hover_css = "fill:#1279BF;stroke:#1279BF;cursor:pointer;")
# }

# view_lmtd_eng_dotplot_int()
```

### Maps

#### Seattle Tracts
This map shows the proportion of pocome households for each census tract in Seattle.
```{r census-geos-tr-map}

tr <- 
        ycc_lmtd_eng_sf %>% 
        filter(GEOGRAPHY == 'tract')

pal <- colorNumeric(viridis(256, option = "C"),tr$LMTD_ENG_PCT_EST)

myLfltGrey() %>% 
        addPolygons(data = as(tr,'Spatial'),
                fillOpacity = .75,
                fillColor = ~pal(LMTD_ENG_PCT_EST),
                color = col2hex('white'),
                opacity = 1,
                weight = .5,
                smoothFactor = 0) %>% 
        addLegend(title = "Limited-English Households",
                  position = 'topright',
                  pal = pal, 
                  values = tr$LMTD_ENG_PCT_EST,
                  opacity = .75,
                  labFormat = labelFormat(suffix = '%',transform = function(x) 100 * x))


```

#### YCC Tract-Neighborhoods
This map reflects the 
```{r census-geos-ycc-nhood-map}

ycc_geoids <- 
        ycc_lmtd_eng_sf %>% 
        filter(!is.na(NAME)) %>% 
        filter(!is.na(GEOID)) %>% 
        select(GEOID) %>% unlist(use.names = FALSE)

tr <- 
        ycc_lmtd_eng_sf %>% 
        filter(GEOGRAPHY == 'tract')

tr_no_ycc <- 
        ycc_lmtd_eng_sf %>% 
        filter(GEOGRAPHY == 'tract') %>% 
        filter(GEOID %!in% ycc_geoids)

nhood <-
        ycc_lmtd_eng_sf %>% 
        filter(GEOGRAPHY == 'neighborhood') 

cnt <- nhood %>% st_union() %>% st_centroid()

lng <- st_bbox(cnt)['xmin'][[1]]

lat <- st_bbox(cnt)['ymin'][[1]]

pal <- colorNumeric(viridis(256, option = "C"),tr$LMTD_ENG_PCT_EST)

myLfltGrey() %>% 
        addPolygons(data = as(tr_no_ycc,'Spatial'),
                fillOpacity = .5,
                fillColor = ~pal(LMTD_ENG_PCT_EST),
                color = col2hex('white'),
                opacity = 1,
                weight = .5,
                smoothFactor = 0) %>% 
         addPolygons(data = as(nhood, 'Spatial'),
                fillColor = ~pal(LMTD_ENG_PCT_EST), 
                fillOpacity = .75,
                color = col2hex('white'),
                opacity = 1,
                weight = .5,
                smoothFactor = 0) %>% 
        setView(lng = lng,lat = lat,zoom = 13) %>% 
        addLegend(title = "Limited-English Households",
                  position = 'topright',
                  pal = pal, 
                  values = tr$LMTD_ENG_PCT_EST,
                  opacity = .75,
                  labFormat = labelFormat(suffix = '%',transform = function(x) 100 * x))
       

```

#### YCC Neighborhood Group
```{r census-geos-ycc-map}

ycc_geoids <- 
        ycc_lmtd_eng_sf %>% 
        filter(!is.na(NAME)) %>% 
        filter(!is.na(GEOID)) %>% 
        select(GEOID) %>% unlist(use.names = FALSE)

tr <- 
        ycc_lmtd_eng_sf %>% 
        filter(GEOGRAPHY == 'tract')

tr_no_ycc <- 
        ycc_lmtd_eng_sf %>% 
        filter(GEOGRAPHY == 'tract') %>% 
        filter(GEOID %!in% ycc_geoids)

ycc <-
        ycc_lmtd_eng_sf %>% 
        filter(GEOGRAPHY == 'neighborhood group') 

cnt <- nhood %>% st_union() %>% st_centroid()

lng <- st_bbox(cnt)['xmin'][[1]]

lat <- st_bbox(cnt)['ymin'][[1]]

pal <- colorNumeric(viridis(256, option = "C"),tr$LMTD_ENG_PCT_EST)

myLfltGrey() %>% 
        addPolygons(data = as(tr_no_ycc,'Spatial'),
                fillOpacity = .5,
                fillColor = ~pal(LMTD_ENG_PCT_EST),
                color = col2hex('white'),
                opacity = 1,
                weight = .5,
                smoothFactor = 0) %>% 
         addPolygons(data = as(ycc, 'Spatial'),
                fillColor = ~pal(LMTD_ENG_PCT_EST), 
                fillOpacity = .75,
                color = col2hex('white'),
                opacity = 1,
                weight = .5,
                smoothFactor = 0) %>% 
        setView(lng = lng,lat = lat,zoom = 13) %>% 
        addLegend(title = "Limited-English Households",
                  position = 'topright',
                  pal = pal, 
                  values = tr$LMTD_ENG_PCT_EST,
                  opacity = .75,
                  labFormat = labelFormat(suffix = '%',transform = function(x) 100 * x))
       

```
